Malacology: A Programmable Storage System Built on Ceph

### 2016-05-10 paper/paper.tex
rel work
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>
Changes to section 1 and 2

\section{Introduction}\label{introduction}

@@

To illustrate the benefits and challenges of this approach we examine
the programmability of Ceph~\cite{weil:osdi2006-ceph}, the increasingly
popular, production-level open-source distributed storage system.
Something of a storage swiss army knife, Ceph supports file, block, and
object interfaces simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous 
Distributed Object Storage (RADOS) system is a cluster of storage
devices (OSDs) that provide Ceph with data durability and integrity
using replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
% By introducing
% programmability concepts into Ceph, we can build new services by
% carefully exposing internal storage services to applications.

@@

To illustrate the benefits and challenges of this approach we have
designed and evaluated Malacology, a programmable storage system for
constructing new services by re-purposing existing subsystem
abstractions contained in Ceph. We build the framework in Ceph by
leveraging a broad spectrum of existing services, including distributed
locking and caching services provided by MDSs, durability and logical
storage devices provided by RADOS, and propogation of consistent cluster
state provided by the monitoring service (see \ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
functionality necessary for implementing new services. Our contributions
are:

\paragraph*{A programmable storage system implementation} that re-uses and extends existing abstractions. It includes:

  \begin{enumerate}
%  \def\labelenumi{\arabic{enumi}.}
%  \itemsep1pt\parskip0pt\parsep0pt
  \item
    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization
  \item
    The use of Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
  \item
    The use of storage objects for persisting policies
  \item
    An extension of the object interface classes for new logical storage/metadata devices
  \end{enumerate}

\paragraph*{Two example implementations and evaluations} for feasibility of services that use this framework

  \begin{enumerate}
%  \def\labelenumi{\arabic{enumi}.}
%  \itemsep1pt\parskip0pt\parsep0pt
  \item
    A high-performance distributed shared log service called Zlog which is based on CORFU~\cite{balakrishnan_corfu_2012}
  \item
    A re-implementation of the programmable Mantle metadata load balancing service~\cite{sevilla:sc15-mantle} using Malacology
  \item 
    An application of the Mantle metadata load balancing service to manage the placement of new sequencer inodes
  \end{enumerate}

First, we describe existing techniques for bridging the gap between the
performance of storage systems and the requirements of applications
(\S\ref{problem}). Next, we discuss the ``distributed systems'' inspired
abstractions employed by Ceph (\S\ref{background}) and allude to their
re-usability as components in other services. The next section
enumerates the abstractions that Malacology exposes in the
implementation section (\S\ref{implementation}). We then show how we use
Malacology to synthesize entirely new storage services on an existing
system through configuration and small changes (\S\ref{services}). We
conclude by evaluating real-world use cases (\S\ref{evaluation}).

\section{Highly Tailored and Application-Specifc Storage
Systems}\label{highly-tailored-and-application-specifc-storage-systems}

@@

A consequence of complicated data management requirements and faster storage
devices is that storage systems may fail to meet the needs of applications.
Application workarounds for meeting performance goals roughly fall into one of
three categories: ``bolt-on'' services, application changes, and changes to
the storage system itself.

\subsection{\texorpdfstring{``Bolt-on''
services}{Bolt-on services}}\label{bolt-on-services}

So called ``bolt-on'' services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on'' services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{balakrishnan_corfu_2012}. Without going into greater detail, the CORFU protocol, by separating the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
``sequencer'' service. This high-performance in-memory counter is used to assign a log
position to each client that appends to the log. The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means). Thus, it is the volatility of the sequencer that enables high performance. A ``bolt-on'' option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the preconceived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

\subsection{Application Changes}\label{application-changes}

The second approach that applications may use to adapt to a storage system
limitation is to extend the responsibility of the application. This means
changing the application itself by adding more data management intelligence or
using domain-specific middleware (e.g., a new data layout). For instance, an
application may be changed to exploit data locality or I/O parallelism in a
distributed storage system.
In SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} both the Hadoop
application and the Hadoop framework itself are changed to leverage the structure of
scientific data (3D array-based data, more specifically) to
improve locality, performance, and increase the number of early results. %This is not a
%bad proposition, but 
However, changing applications creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

@@

When both the bolt-on approach and changing applications fail to meet the needs of the application,
developers turn their attention to the storage system itself.
Traditionally, storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system.
%The
%difficulty of both of these approaches has given rise to a third
%technique called active storage.

@@

As a last resort, the application developer can alter the storage system
itself. In the best case a developer is using an open-source system in which
case one must become familiar with often enormous code-bases before being able
to make changes, but then face the task of approaching a community to
integrate changes, or be forced to maintain an in-house version and risk
maintaining features that become incompatible.  A successful example of this
is the work that focused on HDFS scalability for metadata-intensive workloads~\cite{shvachko_hdfs_2010}. This has lead to modifications to its
architecture or API~\cite{balmin:sigmod2012-clydesdale} to improve
performance.

@@

\subsection{Mantle: A Programmable Metadata Load Balancer}

@@

\subsection{ZLog: A Fast Distributed Shared-Log}

>>>>

\section{Introduction}
\label{introduction}

@@

To illustrate the benefits and challenges of this approach we examine
the programmability of Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
popular, production-level open-source distributed storage system.
Something of a storage swiss army knife, Ceph supports file, block, and
object interfaces simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous 
Distributed Object Storage (RADOS) system is a cluster of storage
devices (OSDs) that provide Ceph with data durability and integrity
using replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
% By introducing
% programmability concepts into Ceph, we can build new services by
% carefully exposing internal storage services to applications.

@@

To illustrate the benefits and challenges of this approach we have
designed and evaluated Malacology, a programmable storage system for
constructing new services by re-purposing existing subsystem
abstractions contained in Ceph. We build the framework in Ceph by
leveraging a broad spectrum of existing services, including distributed
locking and caching services provided by MDSs, durability and logical
storage devices provided by RADOS, and propogation of consistent cluster
state provided by the monitoring service (see Figure~\ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
functionality necessary for implementing new services. Our contributions
are:

\paragraph*{A programmable storage system} that re-uses and extends existing abstractions and services of Ceph, a common production-level storage system. It includes:

  \begin{enumerate}
%  \def\labelenumi{\arabic{enumi}.}
%  \itemsep1pt\parskip0pt\parsep0pt
  \item
    An extension of file system metadata by a new type of ``sequencer'' inode that together with Ceph's capability subsystem implements high-performance, non-durable access serialization:
  \item
    The use of Ceph's PAXOS-based consensus infrastructure for versioning of metadata service load-balancing policies
  \item
    The use of storage objects for persisting policies
  \item
    An extension of the object interface classes for new logical storage/metadata devices
  \end{enumerate}

\paragraph*{A distributed shared log service} to demonstrate the feasibility of services using the programmable storage approach:

  \begin{enumerate}
%  \def\labelenumi{\arabic{enumi}.}
%  \itemsep1pt\parskip0pt\parsep0pt
  \item
    A high-performance distributed shared log service called Zlog which is based on CORFU~\cite{balakrishnan_corfu_2012}
  \item
    A re-implementation of the programmable Mantle metadata load balancing service~\cite{sevilla:sc15-mantle} using Malacology
  \item 
    An application of Mantle to manage load-balancing of sequencers.
  \end{enumerate}

First, we describe existing techniques for adapting storage systems
(\S\ref{problem}). Next, we discuss the ``distributed systems'' inspired
abstractions employed by Ceph (\S\ref{background}) and allude to their
re-usability as components in other services. The next section
enumerates the abstractions that Malacology exposes in the
implementation section (\S\ref{implementation}). We then show how we use
Malacology to synthesize entirely new storage services on an existing
system through configuration and small changes (\S\ref{services}). We
conclude by evaluating real-world use cases (\S\ref{evaluation}).

\section{Adapting Storage Systems}

@@

A consequence of complicated data management requirements and faster storage
devices is that storage systems may fail to meet the needs of applications.
Short of changing the appalication itself, approaches for meeting performance goals roughly fall into ``bolt-on'' services and changes to the storage system itself.

\subsection{Bolt-on services}
\label{bolt-on-services}

Bolt-on services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on'' services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{balakrishnan_corfu_2012}. The CORFU protocol separates the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
``sequencer'' service. This high-performance in-memory counter is used to assign a log
position to each client that appends to the log. The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means). Thus, it is the volatility of the sequencer that enables high performance (see Section~\ref{sec:zlog} for more detail). A ``bolt-on'' option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the preconceived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

\if{false}
\subsection{Application Changes}\label{application-changes}

The second approach that applications may use to adapt to a storage system
limitation is to extend the responsibility of the application. This means
changing the application itself by adding more data management intelligence or
using domain-specific middleware (e.g., a new data layout). For instance, an
application may be changed to exploit data locality or I/O parallelism in a
distributed storage system.
In SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} both the Hadoop
application and the Hadoop framework itself are changed to leverage the structure of
scientific data (3D array-based data, more specifically) to
improve locality, performance, and increase the number of early results. %This is not a
%bad proposition, but 
However, changing applications creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.
\fi

@@

Storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system.
%The
%difficulty of both of these approaches has given rise to a third
%technique called active storage.

@@

To go beyond tuning one has to alter the storage system
itself. In the best case a developer is using an open-source system in which
case one must become familiar with often enormous code-bases before being able
to make changes, but then face the task of approaching a community to
integrate changes, or be forced to maintain an in-house version and risk
maintaining features that become incompatible.  A successful example of this
is the work that focused on HDFS scalability for metadata-intensive workloads~\cite{shvachko_hdfs_2010}. This has lead to modifications to its
architecture or API~\cite{balmin:sigmod2012-clydesdale} to improve
performance.

@@

\subsection{Mantle: A Programmable Metadata Load Balancer}
\label{sec:mantle}

@@

\subsection{ZLog: A Fast Distributed Shared-Log}
\label{sec:zlog}
